A decomposition Approach for solving the JSP using multi-shot ASP\thanks{Supported by organization x.}

### 2022-04-12 ICLP Paper/samplepaper.tex
control algorithm updated

The task of optimizing the horizon of a (partial) schedule means choosing
the execution order of operations of the latest time window sharing a machine
such that the value for the \lstinline{makespan} variable is minimized.
In single-shot ASP modulo DL solving with \clingodl, this can be accomplished
via the command-line option \lstinline{--minimize-variable=makespan}.
\hl{This option, however, is implemented by means of a fixed control loop that
cannot be combined with (other) multi-shot solving methods.}
For the successive optimization by time windows, where the scheduling
horizon gradually increases, we thus require a dedicated treatment of DL constraints
limiting the value for \lstinline{makespan}.
To this end, the \lstinline{optimize(m)} subprogram below line~\ref{prg:diff_log:oper_min}
declares an external atom \lstinline{horizon(m)} for controlling whether a DL
constraint asserted in line~\ref{prg:diff_log:oper_min:end} is active and limits
the \lstinline{makespan} value to an integer supplied for the parameter~\lstinline{m}.
When launching the optimization process for a new time window, any instance of the
\lstinline{horizon(m)} atom introduced before is set to false for making sure that
some (partial) schedule is feasible.
Once some schedule with a horizon $h+1$ is found,
the subprogram \lstinline{optimize(}$h$\lstinline{)} is instantiated on demand, i.e.,
in case $h$ has not been passed as a value for~\lstinline{m} before, and
the corresponding \lstinline{horizon(}$h$\lstinline{)} atom is set to true for activating the DL
constraint reducing the admitted scheduling horizon to~$h$.
The process of successively reducing the horizon~$h$ in order to find better partial schedules
is repeated until the imposed \lstinline{makespan} value turns out to be infeasible,
meaning that an optimal partial schedule has been found.
As already mentioned above, the introduced instances of the external \lstinline{horizon(m)} atom
are then set to false and the successive optimization proceeds by instantiating the
\lstinline{step(}$w+1$\lstinline{)} subprogram for the next time window $w+1$ (if any), and
also supplying the determined starting times of operations from time window~$w$
by reified facts.
The described control loop for successively extending good-quality partial schedules to
a global solution is implemented by means of the Python interface of \clingodl,
where a time limit can be configured to restrict the optimization efforts per time window
and thus make sure that the iterative scheduling progresses.
\hl{Figure~{\ref{fig:enum}} summarizes the main steps of the Multi-shot solving controlled by a python script.}

\begin{figure}[tb]
\begin{tcolorbox}[breakable,arc=0mm,outer arc=0mm,colbacktitle=lightgray!50!white,coltitle=black,colframe=gray,fonttitle=\bfseries,colback=white,subtitle style={colback=white}]
%  \vspace{0.1\baselineskip}
Starting with $H:=\emptyset$, for $w:= 1$ to number $n$ of time windows, do:
\begin{enumerate}[label=\textbf{\arabic*.}]
  \item % \begin{tabbing}
        If $w= 1$, % \= 
        then $P:=\{\text{\lstinline{base}}\}$;
        % \\
        % \>

        % \hfill
        else $P := \{\text{\lstinline{start((}$j$\lstinline{,}$s$\lstinline{),}$t$\lstinline{,}$w-1$\lstinline{).}}
                    \mid
                    \{\text{\lstinline{window(}$j$\lstinline{,}$s$\lstinline{,}$w-1$\lstinline{)}},
                    \text{\lstinline{(}$j$\lstinline{,}$s$\lstinline{)}}=t\}\subseteq X\}$.
        % \end{tabbing}
  \item Ground $P\cup\{\text{\lstinline{step(}$w$\lstinline{)}}\}$
  \item While the time limit per time window is not reached, do:
  \begin{enumerate}[label=\textbf{\alph*.}]
  \item Find some answer set $X$
  \end{enumerate}
  \item Apply the compression of a window (postprocessing phase).
  \item Keep the operations assignment to be passed to the next window.
  \item Go to \textbf{2} until all windows are optimized.
\end{enumerate}
  \begin{enumerate}[label=\textbf{\arabic*.}]
    \item Ground the base program.
    \item Ground the next shot.
    \item Optimize the current window till reaching the optima or the timeout.
    \item Apply the compression of a window (postprocessing phase).
    \item Keep the operations assignment to be passed to the next window.
    \item Go to \textbf{2} until all windows are optimized.
  \end{enumerate}
\end{tcolorbox}
\caption{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
\end{figure}
% \captionof{figure}{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}

>>>>

\begin{figure}[tb]
  \begin{tcolorbox}[breakable,arc=0mm,outer arc=0mm,colbacktitle=lightgray!50!white,coltitle=black,colframe=gray,fonttitle=\bfseries,colback=white,subtitle style={colback=white}]
  %  \vspace{0.1\baselineskip}
  Starting with $H:=\emptyset$, for $w:= 1$ to number $n$ of time windows:
  \begin{enumerate}[label=\textbf{\arabic*.}]
    \item For each $h\in H$, set \lstinline{horizon(}$h$\lstinline{)} to false
    \item % \begin{tabbing}
          If $w= 1$, % \= 
          then $P:=\{\text{\lstinline{base}}\}$;
          % \\
          % \>

          % \hfill
          else $P := \{\text{\lstinline{start((}$j$\lstinline{,}$s$\lstinline{),}$t$\lstinline{,}$w-1$\lstinline{).}}
                      \mid
                      \{\text{\lstinline{window(}$j$\lstinline{,}$s$\lstinline{,}$w-1$\lstinline{)}},
                      \text{\lstinline{(}$j$\lstinline{,}$s$\lstinline{)}}=t\}\subseteq X\}$
          % \end{tabbing}
    \item Ground $P\cup\{\text{\lstinline{step(}$w$\lstinline{)}}\}$
    \item While there is some answer set $X$ and the time limit per time window is not reached:
      \begin{enumerate}[label=\textbf{\alph*.}]
      \item $h:=t-1$, where $($\lstinline{makespan}${}=t)\in X$
      \item If $h\notin H$, then $H := H\cup\{h\}$ and ground $\{$\lstinline{optimize(}$h$\lstinline{)}$\}$
      \item Set \lstinline{horizon(}$h$\lstinline{)} to true
      \end{enumerate}
    \end{enumerate}
  \end{tcolorbox}
  \caption{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
  \end{figure}
  % \captionof{figure}{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
%
The task of optimizing the horizon of a (partial) schedule means choosing
the execution order of operations of the latest time window sharing a machine
such that the value for the \lstinline{makespan} variable is minimized.
In single-shot ASP modulo DL solving with \clingodl, this can be accomplished
via the command-line option \lstinline{--minimize-variable=makespan}.
\hl{This option, however, is implemented by means of a fixed control loop that
cannot be combined with (other) multi-shot solving methods.}
For the successive optimization by time windows, where the scheduling
horizon gradually increases, we thus require a dedicated treatment of DL constraints
limiting the value for \lstinline{makespan}.
To this end, the \lstinline{optimize(m)} subprogram below line~\ref{prg:diff_log:oper_min}
declares an external atom \lstinline{horizon(m)} for controlling whether a DL
constraint asserted in line~\ref{prg:diff_log:oper_min:end} is active and limits
the \lstinline{makespan} value to an integer supplied for the parameter~\lstinline{m}.
When launching the optimization process for a new time window, any instance of the
\lstinline{horizon(m)} atom introduced before is set to false for making sure that
some (partial) schedule is feasible.
Once some schedule with a horizon $h+1$ is found,
the subprogram \lstinline{optimize(}$h$\lstinline{)} is instantiated on demand, i.e.,
in case $h$ has not been passed as a value for~\lstinline{m} before, and
the corresponding \lstinline{horizon(}$h$\lstinline{)} atom is set to true for activating the DL
constraint reducing the admitted scheduling horizon to~$h$.
The process of successively reducing the horizon~$h$ in order to find better partial schedules
is repeated until the imposed \lstinline{makespan} value turns out to be infeasible,
meaning that an optimal partial schedule has been found.
As already mentioned above, the introduced instances of the external \lstinline{horizon(m)} atom
are then set to false and the successive optimization proceeds by instantiating the
\lstinline{step(}$w+1$\lstinline{)} subprogram for the next time window $w+1$ (if any), and
also supplying the determined starting times of operations from time window~$w$
by reified facts.
The described control loop for successively extending good-quality partial schedules to
a global solution is implemented by means of the Python interface of \clingodl,
where a time limit can be configured to restrict the optimization efforts per time window
and thus make sure that the iterative scheduling progresses.
\hl{Figure~{\ref{fig:enum}} summarizes the main steps of the Multi-shot solving controlled by a python script.}
